---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# iadf
Calculate false ring proportions 
  from data frames of intra annual density fluctuations.
  
If you know any further methods which should get included or find any bugs or 
mistakes in the code, please contact the package maintainer or file an issue on GitHub. 

## Installation
`iadf` is available on CRAN and on GitHub:

To install the stable version from CRAN type:
```{r CRAN-installation, eval = FALSE}
install.packages("iadf")
```


You can install the latest development version of the package from GitHub with:

```{r gh-installation, eval = FALSE}
# install.packages("devtools")
devtools::install_github("konradmayer/iadf")
```

To also install the vignettes type:
```{r gh-vignette, eval = FALSE}
# install.packages("devtools")
devtools::install_github("konradmayer/iadf", build_vignettes = TRUE)
```


## Updating

Subsequent implementations and bug fixes require regular updating, to keep your
installation up to date type:

```{r gh-update, eval = FALSE}
devtools::update_packages("iadf")
```
