---
title: "False-ring proportion"
author: "Konrad Mayer"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
bibliography: bibliography.bib
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

# Introduction
Intra annual density fluctuations (iadfs), also referred as false rings are 
defined as either latewood-like cells in earlywood or earlywood-like cells in
latewood [@Fritts1976]. Although recent efforts are made to specify the 
position, extend as well as intensity of iadfs most studies use binary 
assignments to indicate presence or absence of iadfs [@Battipaglia2016]. While
frequencies can easily be calculated as the proportion of rings showing an 
iadf in a specific year several studies suggest the consideration of variing
variances, age trends as well as influence of ring size. This package implements 
all published approaches known by the package author, their usage is shortly
explained in this vignette.

# Data
The data used in this vignette is included in the package. You can load the
data-sets called "example_rwl" and "example_iadf" by typing:

```{r}
library("iadf")
data("example_rwl")
data("example_iadf")
```

example_rwl is a data frame, similar to the rwl class defined by the package 
`dplR` with series as columns and years as rows, rownames specifying the year.
The data shows the ring width with years not covered by the sample marked with 
`NA`.

```{r, echo=FALSE, results='asis'}
knitr::kable(example_rwl[20:30, 10:16])
```

example_iadf has the same dimensions as example_rwl, just showing binary
assignements of iadfs instead of ring width.

```{r, echo=FALSE, results='asis'}
knitr::kable(example_iadf[20:30, 10:16])
```


# False-ring proportions

## Naive approach
The proportion of rings showing an iadf can easily be calculated using base R 
functions:

```{r, results='hide'}
rowMeans(example_iadf, na.rm = TRUE)
```

However, there's also a function in the package `iadf` to calculate the false
ring proportion with data frame output consistent to the other package functions
and improved warning messages. 

```{r, results='hide'}
(results_frp <- frp(example_iadf))
```

Please mention that this function is somehow slower 
than using `rowMeans()`, so consider using the base R code whenever computation
speed is critical.

## Consideration of sample size
As the variance of time series is dependend on sample size the variance can be
adjusted according to Osborn [-@Osborn1997], using the function `afrp()`:

```{r, results='hide'}
(results_afrp <- afrp(example_iadf))
```

## Consideration of age trend 
As other tree ring parameters, also IADF occurence shows an age trend. 
Novak [-@Novak2013] suggested a detrending procedure to reduce this bias.

First the iadf frequency per cambial age needs to be calculated:
```{r, results='hide'}
frq <- novak_freq(example_iadf)
```

Then we try to model the influence of age on iadf frequency using a Weibull
function as suggested by Novak [-@Novak2013], limiting the data pairs used to
only frequencies at cambial ages calculated from at least 15 years:

```{r, error=TRUE, warning=TRUE}
mdl <- novak_weibull(frq, 15)
```
Which results in an error. This can either be due to insufficient starting values 
for the curve fitting function (which can be solved using
`novak_negexp_find_start()` or `novak_weibull_find_start()` and will be discussed
in the next section) or due to a wrong curve 
type. Regardless the suggestion of Novak [-@Novak2013] to fit a Weibull function,
we found negative exponential curves fit much better to our data (maybe because
of differences in iadf formation due to the different climatic regime as 
Novak [-@Novak2013] uses data from mediterranean sites while the presented data
comes from temperate climate)

Therefore we can use

```{r, fig.width=7}
mdl <- novak_negexp(frq, 15)
```

to fit the curve. Currently there are only functions for fitting Weibull and 
negative exponential curves available in the package. If other curve types fit
your data better, please contact the package maintainer to include this type as
well.

Next we hand the model and the original data to the function `novak_index()`:
```{r, results='hide'}
results_novak <- novak_index(example_iadf, mdl)
```

to calculate the iadf proportion with age trend removed.

## Consideration of the influence of ring width 
Campelo [-@Campelo2014] states that beneath ring age also ring with influences
iadf formation and introduced another standardization approach.

The workflow is almost the same as for the approach above. Fist we calculate 
frequencies per ring width class using both data sets:

```{r}
frq <- campelo_freq(example_iadf, example_rwl)
```

Then we fit a chapman function to our frequencies:

```{r, fig.width=7}
mdl <- campelo_chapman(frq)
```

In case the function throws an error it's likely due to insufficient starting
values. These can be found interactively with `campelo_chapman_find_start()` and
then used in `campelo_chapman()`:

```{r, fig.width=7, eval=FALSE}
st <- campelo_chapman_find_start(frq)
mdl <- campelo_chapman(frq, start = st)
```

Next the index is calculated using both data sets and the model:

```{r, results='hide'}
results_campelo <- campelo_index(example_iadf, example_rwl, mdl)
```

#Comparison of the approaches

```{r, fig.width=7, fig.height=5}
plot(NULL, xlim = range(as.numeric(rownames(example_iadf))), 
     ylim = c(-0.5, 2.5), xlab = '', ylab = '')
lines(results_frp, col = 'blue')
lines(results_afrp, col = 'green')
lines(results_novak, col = 'purple')
lines(results_campelo, col = 'red')
legend('topright', col = c('blue', 'green', 'purple', 'red'), 
       legend = c('frp', 'afrp', 'Novak', 'Campelo'), bty = 'n', lty = 1)
```



#References

